Anomalous waves triggered by abrupt depth changes: experiments and truncated KdV theory

### 2019-10-15 paper2_JNLS/paper.tex
NM

\begin{abstract} 
Recent laboratory experiments revealed the emergence of anomalous statistics in a randomized surface waves upon encountering an abrupt depth change (ADC) in the bottom topography (Bolles et al. PRF 2019). Majda et al.~2019 established a theoretical foundation, based on analysis of the truncated KdV (TKdV) system, that successfully captures several key qualitative features of the experiments including the robust emergence of anomalous statistics. The framework exploits the Hamiltonian structure of TKdV to construct invariant measures that are statistical matched at the ADC, and is corroborated by simulations of the deterministic dynamics. 
%
Here we present a more comprehensive treatment of these parallel experimental and modeling efforts, with emphasis on the link between the two. We present an assortment of new experimental measurements and carefully compare them to model predictions. We establish a more precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in experiments. The model predicts a peculiar relationship between surface-displacement skewness and surface-slope variance, which is confirmed in spectacular fashion by direct experimental measurements.
\end{abstract}

@@

The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al~\cite{bolles2019} and Majda et al.~\cite{majda2019, majdaqi2019}. Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}. We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory. Likewise, novel theoretical predictions have inspired us to perform additional experimental measurements, especially as it pertains to a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory. Comparison with the new measurements further confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}. In addition, we provide a more thorough treatment of the link between experimental and model parameters to facilitate comparison between the two. In particular, we flesh out details of non-dimensionalization, which were only briefly addressed in prior work for the sake of brevity.

@@

The pivoting motion of the paddle is driven by a 5-phase stepper motor. To generate a randomized wave field, the paddle angle $\phi$ is specified by a psuedo-random signal
\begin{align}
\label{PaddleAngle}
& \phi(t) = \phi_0 + \Dphi \sum_{n=1}^N a_n \cos(\omega_n t+\delta_n) \, , \\
\label{anEq}
& a_n = \sqrt{\frac{2 \Delta \omega}{\pi^{1/2} \omsig}} \, 
\exp \left( -\frac{(\omega_n - \omavg)^2}{2 \omsig^2} \right) \, ,
\end{align}
The angular frequencies are evenly spaced $\omega_n = n  \Delta \omega$ with step size $ \Delta \omega = (\omavg+4 \omsig)/N$, where $\omavg$ and $\omsig$ represent the mean and the bandwidth of $\omega$ respectively. As in prior work, all experiments reported here use the values $\omavg = \omsig = 12.5$ rad/s, corresponding to a peak forcing frequency of 2 Hz and bandwidth of 2 Hz. The phases $\delta_n$ are uniformly distributed random variables, which results in a randomized wave train. The standard deviation of the paddle angle, $\Dphi$, controls the amplitude overall amplitude of the waves. In a single experiment $\Dphi$ is fixed, and we will present a series of experiments with $\Dphi$ varied systematically.

@@

	Example measurements of free-surface displacements $\eta$ are shown in Figs.~\ref{ExpDiagStats}(b)--(c). These measurements are extracted from the images at two representative locations: one a short distance (9 cm) upstream of the ADC and the other a short distance (15 cm) downstream. Both signals exhibit a combination of periodic and random behavior, with the dominant oscillations corresponding to the peak forcing frequency of 2 Hz. The nature of the random fluctuations is revealed by the corresponding histograms shown in Fig.~\ref{ExpDiagStats}(d)--(e) on a semi-log scale. The upstream measurements are symmetrically distributed about the mean, $\eta = 0$. In fact, Bolles et al.~found that these measurements follow a Gaussian distribution closely \cite{bolles2019}. The downstream measurements, however, skew strongly towards positive displacement, $\eta > 0$. Bolles et al.~found these measurements to be well described by a mean-zero gamma distribution \cite{bolles2019}. The slower decay rate of this downstream distribution indicates an elevated level of extreme surface displacement, i.e.~rogue waves. Bolles et al.~estimated that a rogue wave can be up to 65 times more likely in these experiments than if displacements were Gaussian \cite{bolles2019}.

% Figure: Experimental Spatial Statistics
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.85 \linewidth]{Figs/ExpSpatialStats.pdf}
\caption{Wave statistics as they vary in space for several different experiments of different driving amplitudes (see legend). (a)--(c) Standard deviation, skewness, and kurtosis of the surface displacement $\eta$. (d)--(f) The same for surface slope $\eta_x$. A reference wave amplitude is set by $\etastd$ which does not vary significantly crossing the ADC. Skewness and kurtosis of $\eta$, however, show a strong response to the ADC. The surface slope measurements are negatively skewed and also show a large kurtosis downstream of the ADC. }
\label{ExpSpatialStats}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

	The paddle amplitude in Fig.~\ref{ExpDiagStats} is $\Dphi = 1.38^{\circ}$, and this value was varied systematically in the range $\Dphi = 0.125^{\circ}$--$2^{\circ}$ to probe the different regimes of wave behavior, from linear to strongly nonlinear waves. Figure \ref{ExpSpatialStats} shows long-time statistics of both surface displacement $\eta$ and the surface slope $\eta_x$ as they vary in space for six different driving amplitudes (see legend). We will only be examining statistics of mean-zero quantities in this paper, and so, for an arbitrary mean-zero quantity $q$, we have the following definitions
\begin{align}
%& \var(q) = \mean{q^2}		&&\mbox{\em variance} \\
& \std(q) = q_{std}= \sqrt{\mean{q^2}}
&&\mbox{\em standard deviation} \\
& \skw(q) = {\mean{q^3}} / {q_{std}^3}	
&&\mbox{\em skewness} \\
& \kurt(q) = {\mean{q^4}} / {q_{std}^4} - 3
&&\mbox{\em (excess) kurtosis}
\end{align}
where $\mean{}$ indicates a mean (here a long-time mean). Hereafter, we will simply refer to the excess kurtosis as kurtosis.

	Figure \ref{ExpSpatialStats} shows how these statistics vary in the vicinity of the ADC, located at $x = 0$, for both displacement and slope. First, the standard deviation of displacement, $\etastd$, gives the coarsest possible estimate for the amplitude of waves. Figure \ref{ExpSpatialStats}(a) shows that while $\etastd$ increases with driving amplitude, it remains relatively uniform in space for each individual experiment, indicating that the overall amplitude of the wave train is not significantly altered by the presence of the ADC. In experiments with sufficiently high driving amplitude, though, the skewness and kurtosis (b)--(c) respond strongly to the ADC. In these experiments, both the skewness and kurtosis are relatively small upstream of the ADC, indicating nearly Gaussian statistics, but then increase dramatically downstream and reach a peak near $x = 15$ cm. Remarkably, the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude. The maximum values of skewness and kurtosis (roughly 0.9 and 0.7 respectively) seen in the figure indicate a significant departure from Gaussian statistics, as is constant with the histogram in Fig.~\ref{ExpDiagStats}(e). Bolles et al.~found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution. We note that we have gathered data for 15 different driving amplitudes, but only display 6 in Fig.~\ref{ExpSpatialStats} to avoid clutter.

	To complement the displacement statistics, we report here statistics for the surface slope $\eta_x$ in the right column of Fig.~\ref{ExpSpatialStats}. Our attention to slope statistics was motivated by new theoretical developments by Majda et al.~\cite{majda2019} and will be expanded upon in later sections. We extract the surface slope by numerical differentiating images of the free surface using Savitzky-Golay smoothing filters. This ability to extract surface slope is another advantage of our optical measurements over the more commonly used technique of placing a set of discrete wave probes in the tank. As seen in Fig.~\ref{ExpSpatialStats}(a) the standard deviation of slope behaves similar to displacement: $\std(\eta_x)$ increases with driving amplitude, but remains nearly uniform in space for each individual experiment and is not affected by the ADC. The higher-order moments, however, once again respond strongly to the ADC if the driving amplitude is sufficiently high. The skewness of slope becomes highly negative and reaches a minimum near $x = 8$ cm, which is downstream of the ADC but upstream of the position where the displacement statistics peak. The kurtosis of $\eta_x$ reaches a peak at nearly the same location as skewness. We note that the large {\em negative} skewness of $\eta_x$ indicates a bias towards negative slope, which is consistent with a right-moving wave of steep leading surface and shallower trailing surface, i.e.~a wave that is near overturning.

@@

We now introduce the theoretical framework that will be used to understand and quantify the experimental observations. This framework is based on a Galerkin truncation of the variable-depth Korteweg–de Vries (KdV) equation. The KdV equation is a commonly excepted model for describing the propagation of unidirectional, shallow-water waves, and it accounts for weak nonlinearity and weak dispersion over long timescales and large spatial scales. We will perform Galerkin truncation of this system at wavenumber $\Lambda$ to obtain a finite-dimensional dynamical system that exhibits weak turbulence. We outline the Hamiltonian structure of both the continuous KdV and the truncated system. This structure is exploited to obtain invariant measures of the underlying dynamics and, ultimately, to rationalize the experimental findings on anomalous wave statistics triggered by an ADC.

@@

In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form chosen for convenience in working with the statistical-mechanics framework introduced in \cite{majda2019}. Since the choice of normalization not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other choices. To this end, consider characteristic scales, $\ampscale, \lengthscale, \timescale$ for the wave amplitude, longitudinal length, and time respectively, which can remain unspecified for the moment. We introduce the dimensionless variables
\begin{align}
&u = \eta / \ampscale
&&\mbox{\em dimensionless surface displacement} \\
&\tilde{x} = \xi / \lengthscale
&&\mbox{\em dimensionless position (in moving frame)} \\
&\tilde{t} = t / \timescale
&&\mbox{\em dimensionless time}
\end{align}
Recasting \eqref{KdV} in terms of these variables gives the generic dimensionless KdV equation:
\begin{equation}
u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
+ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
\end{equation}
We have dropped the tilde notation above for simplicity and will henceforth use tildes only in cases of possible ambiguity.

Now it is possible to choose the scales $\ampscale, \lengthscale, \timescale$ for ease in working with a particular framework. We choose the following scales,
\begin{align}
\label{scales}
\ampscale = \pi^{1/2} \, \etastd \, , \qquad
\lengthscale_{\pm} = \frac{\lamfac \lamupdn}{2 \pi} \, , \qquad
\timescale_{\pm} = \frac{\lamfac \lamupdn}{2 \pi \freqp \dupdn}
\end{align}
where $\lamfac$ is an integer to be chosen later. 
The explanation for these choices is as follows. First, we have chosen the characteristic amplitude, $\ampscale$, above to normalize the energy of the state-variable $u$ to unity, as will be demonstrated in Section \ref{tKdVSec}. Second, regarding $\lengthscale$, recall that $\lam$ is the characteristic wavelength corresponding to the peak forcing frequency $\freqp$ in the experiments. If only integer multiples of $\freqp$ were imposed (e.g.~lower frequencies were not present), then the forcing would produce waves that are periodic over lengthscale $\lam$. Since lower frequencies do exist, strict periodicity is not satisfied, but rather waves may be nearly periodic over the physical domain $\xi \in [-\lam/2, \lam/2]$. The approximation of near-periodicity becomes more accurate if integer multiples of $\lam$ are considered, i.e.~$\xi \in [-\lamfac \lam/2, \lamfac \lam/2]$. Thus, we have chosen $\lengthscale$ above so that the dimensionless domain of consideration is $\tilde{x} \in [-\pi, \pi]$ and periodic boundary conditions on $u$ can be  imposed on over this domain with accuracy that increases with $\lamfac$. Lastly, regarding $\timescale$, the most basic timescale in the experiments is simply $\freqp^{-1}$, i.e.~the period of waves passing a fixed reference point. Of course, the leading-order behavior in shallow water is simple translation of waves with uniform speed $c$. The KdV equation provides the first correction to this behavior and describes dynamics that evolve over longer timescales. Hence we have rescaled $\freqp^{-1}$ by the factor $N \lam/(2 \pi \depth) \gg 1$, which provides a suitable long timescale in line with previous normalizations \cite{johnson1997modern}. The scales $\lengthscale = \lengthscale_{\pm}$ and $\timescale = \timescale_{\pm}$ change value across the ADC, which is important to note when comparing the theory against experimental measurements.

>>>>

\begin{abstract} 
Recent laboratory experiments revealed the emergence of anomalous statistics in a randomized surface waves upon encountering an abrupt depth change (ADC) in the bottom topography (Bolles et al. PRF 2019). Majda et al.~2019 established a theoretical foundation, based on analysis of the truncated KdV (TKdV) system, that successfully captures several key qualitative features of the experiments including the robust emergence of anomalous statistics. The framework exploits the Hamiltonian structure of TKdV to construct invariant measures that are statistical matched at the ADC, and is corroborated by simulations of the deterministic dynamics. 
%
Here we present a more comprehensive treatment of these parallel experimental and modeling efforts, with emphasis on the link between the two. We present an assortment of new experimental measurements and carefully compare them to model predictions. We establish a precise link between experimental and model parameters, so that the invariant measure underlying the model can be calibrated to the specific conditions present in experiments. The model predicts a peculiar relationship between surface-displacement skewness and surface-slope variance, which is confirmed in spectacular fashion by direct experimental measurements.
\end{abstract}

@@

The purpose of the present manuscript is to provide a more comprehensive treatment of the {\em combined} experimental and theoretical efforts of Bolles et al~\cite{bolles2019} and Majda et al.~\cite{majda2019, majdaqi2019}. Special emphasis is placed on the synergy between experiment and theory -- a cooperative strategy with a proven record of success \cite{camassa2012stratified, ristroph2012, ganedi2018equilibrium}. 
%We present an assortment of previously unpublished experimental data, for example on surface slope statistics and autocorrelation, which motivates further examination and refinement of the theory. 
Novel theoretical predictions have inspired us to perform additional experimental measurements, especially as it pertains to a peculiar relationship between surface-displacement skewness and surface-slope variance that is predicted by the theory. Comparison with the new measurements further confirms the predictive power of the framework developed by Majda et al.~\cite{majda2019}. In addition, we provide a more thorough treatment of the link between experimental and model parameters to facilitate comparison between the two. In particular, we flesh out details of non-dimensionalization, which were only briefly addressed in prior work for the sake of brevity.

@@

The pivoting motion of the paddle is driven by a 5-phase stepper motor. To generate a randomized wave field, the paddle angle $\phi$ is specified by a psuedo-random signal
\begin{align}
\label{PaddleAngle}
& \phi(t) = \phi_0 + \Dphi \sum_{n=1}^N a_n \cos(\omega_n t+\delta_n) \, , \\
\label{anEq}
& a_n = \sqrt{\frac{2 \Delta \omega}{\pi^{1/2} \omsig}} \, 
\exp \left( -\frac{(\omega_n - \omavg)^2}{2 \omsig^2} \right) \, ,
\end{align}
The angular frequencies are evenly spaced $\omega_n = n  \Delta \omega$ with step size $ \Delta \omega = (\omavg+4 \omsig)/N$, where $\omavg$ and $\omsig$ represent the mean and the bandwidth of $\omega$ respectively. As in prior work, all experiments reported here use the values $\omavg = \omsig = 12.5$ rad/s, corresponding to a peak forcing frequency of 2 Hz and bandwidth of 2 Hz. The phases $\delta_n$ are uniformly distributed random variables, which results in a randomized wave train. The standard deviation of the paddle angle, $\Dphi$, controls the overall amplitude of the waves. In a single experiment $\Dphi$ is fixed, and we will present a series of experiments with $\Dphi$ varied systematically.

@@

	Example measurements of free-surface displacements $\eta$ are shown in Figs.~\ref{ExpDiagStats}(b)--(c). These measurements are extracted from the images at two representative locations: one a short distance (9 cm) upstream of the ADC and the other a short distance (15 cm) downstream. Both signals exhibit a combination of periodic and random behavior, with the dominant oscillations corresponding to the peak forcing frequency of 2 Hz.

	The nature of the random fluctuations is revealed by the corresponding histograms shown in Fig.~\ref{ExpDiagStats}(d)--(e) on a semi-log scale. The upstream measurements are symmetrically distributed about the mean, $\eta = 0$. In fact, Bolles et al.~found that these measurements follow a Gaussian distribution closely \cite{bolles2019}. The downstream measurements, however, skew strongly towards positive displacement, $\eta > 0$. Bolles et al.~found these measurements to be well described by a mean-zero gamma distribution \cite{bolles2019}. The slower decay rate of this downstream distribution indicates an elevated level of extreme surface displacement, i.e.~rogue waves. Bolles et al.~estimated that a rogue wave can be up to 65 times more likely in these experiments than if displacements were Gaussian \cite{bolles2019}.

	The paddle amplitude in Fig.~\ref{ExpDiagStats} is $\Dphi = 1.38^{\circ}$, and this value was varied systematically in the range $\Dphi = 0.125^{\circ}$--$2^{\circ}$ to probe the different regimes of wave behavior, from linear to strongly nonlinear waves. Figure \ref{ExpSpatialStats} shows long-time statistics of both surface displacement $\eta$ and the surface slope $\eta_x$ as they vary in space for six different driving amplitudes (see legend).

	We will only be examining statistics of mean-zero quantities in this paper, and so, for an arbitrary mean-zero quantity $q$, we have the following definitions
\begin{align}
%& \var(q) = \mean{q^2}		&&\mbox{\em variance} \\
& \std(q) = q_{std}= \sqrt{\mean{q^2}}
&&\mbox{\em standard deviation} \\
& \skw(q) = {\mean{q^3}} / {q_{std}^3}	
&&\mbox{\em skewness} \\
& \kurt(q) = {\mean{q^4}} / {q_{std}^4} - 3
&&\mbox{\em (excess) kurtosis}
\end{align}
where $\mean{}$ indicates a mean (here a long-time mean). Hereafter, we will simply refer to the excess kurtosis as kurtosis.

% Figure: Experimental Spatial Statistics
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{figure}%[!ht]
\begin{center}
\includegraphics[width = 0.80 \linewidth]{Figs/ExpSpatialStats.pdf}
\caption{Wave statistics as they vary in space for several experiments of different driving amplitudes (see legend). (a)--(c) Standard deviation, skewness, and kurtosis of the surface displacement $\eta$. (d)--(f) The same for surface slope $\eta_x$. (a) $\etastd$, which sets a scale for wave amplitude, does not vary significantly crossing the ADC. (b)--(c) Skewness and kurtosis of $\eta$, however, show a strong response to the ADC. The surface slope measurements are negatively skewed and also show a large kurtosis downstream of the ADC. }
\label{ExpSpatialStats}
\end{center}
\end{figure}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%

	Figure \ref{ExpSpatialStats} shows how these statistics vary in the vicinity of the ADC, located at $x = 0$, for both displacement and slope. First, the standard deviation of displacement, $\etastd$, gives the coarsest possible estimate for the amplitude of waves. Figure \ref{ExpSpatialStats}(a) shows that while $\etastd$ increases with driving amplitude, it remains relatively uniform in space for each individual experiment, indicating that the overall amplitude of the wave train is not significantly altered by the presence of the ADC. The skewness and kurtosis, however, respond strongly to the ADC as long as the amplitude is sufficiently high. As seen in \ref{ExpSpatialStats}(b)--(c), both the skewness and kurtosis are relatively small upstream of the ADC, indicating nearly Gaussian statistics, but then increase dramatically downstream and reach a peak near $x = 15$ cm. Remarkably, the location of the peak is the same for skewness and kurtosis and appears insensitive to driving amplitude. The maximum values of skewness and kurtosis (roughly 0.9 and 0.7 respectively) seen in the figure indicate a significant departure from Gaussian statistics, as is constant with the histogram in Fig.~\ref{ExpDiagStats}(e). Bolles et al.~found that, once a threshold driving amplitude is exceeded (roughly $\Dphi = 0.5^{\circ}$), the displacement statistics in this anomalous region are robustly described by the gamma distribution. We note that we have gathered data for 15 different driving amplitudes, but only display 6 in Fig.~\ref{ExpSpatialStats} to avoid clutter.

	To complement the displacement statistics, we report here statistics for the surface slope $\eta_x$ in the right column of Fig.~\ref{ExpSpatialStats}. Our attention to slope statistics was motivated by new theoretical developments by Majda et al.~\cite{majda2019} and will be expanded upon in later sections. We extract the surface slope by numerical differentiating images of the free surface using Savitzky-Golay smoothing filters. This ability to extract surface slope is another advantage of our optical measurements over the more commonly used technique of placing a set of discrete wave probes in the tank. As seen in Fig.~\ref{ExpSpatialStats}(a) the standard deviation of slope behaves similar to displacement: $\std(\eta_x)$ increases with driving amplitude, but remains nearly uniform in space for each individual experiment and is not affected by the ADC. Once again though, the higher-order moments respond strongly to the ADC if the driving amplitude is sufficiently high. The skewness of slope becomes highly negative and reaches a minimum near $x = 8$ cm, which is downstream of the ADC but upstream of the position where the displacement statistics peak. The kurtosis of $\eta_x$ reaches a peak at nearly the same location as skewness. We note that the large {\em negative} skewness of $\eta_x$ indicates a bias towards negative slope, which is consistent with a right-moving wave of steep leading surface and shallower trailing surface, i.e.~a wave that is near overturning.

@@

We now introduce the theoretical framework that will be used to understand and quantify the experimental observations. This framework is based on a Galerkin truncation of the variable-depth Korteweg–de Vries (KdV) equation. The KdV equation is a well established model for describing the propagation of unidirectional, shallow-water waves, and it accounts for weak nonlinearity and weak dispersion over long timescales and large spatial scales. We will perform Galerkin truncation of KdV to obtain a finite-dimensional dynamical system that exhibits weak turbulence. We outline the Hamiltonian structure of both the continuous KdV and the truncated system. This structure is exploited to obtain invariant measures of the underlying dynamics and, ultimately, to rationalize the experimental findings on anomalous wave statistics triggered by an ADC.

@@

In this section, the variable-depth KdV equation \eqref{KdV} will be recast into a dimensionless form that is chosen for convenience in working with the statistical-mechanics framework from Majda et al.~2019 \cite{majda2019}. Since the choice of normalization not unique, it is instructive to first introduce a generic normalization to facilitate comparison with other possible choices. To this end, consider characteristic scales, $\ampscale, \lengthscale, \timescale$ for the wave amplitude, longitudinal length, and time respectively, which can remain unspecified for the moment. We introduce the dimensionless variables
\begin{align}
&u = \eta / \ampscale
&&\mbox{\em dimensionless surface displacement} \\
&\tilde{x} = (x-ct) / \lengthscale
&&\mbox{\em dimensionless position (in moving frame)} \\
&\tilde{t} = t / \timescale
&&\mbox{\em dimensionless time}
\end{align}
Recasting \eqref{KdV} in terms of these variables gives the generic dimensionless KdV equation:
\begin{equation}
u_t + \frac{3}{2} \left( \frac{c \timescale \ampscale}{\lengthscale \depth} \right) u u_x 
+ \frac{1}{6} \left( \frac{c \timescale \depth^2}{\lengthscale^3} \right) u_{xxx} = 0
\end{equation}
We have dropped the tilde notation above for simplicity and will henceforth use tildes only in cases of possible ambiguity.

Now it is possible to choose the scales $\ampscale, \lengthscale, \timescale$ for ease in working with a particular framework. We make the following choices,
\begin{align}
\label{scales}
\ampscale = \pi^{1/2} \, \etastd \, , \qquad
\lengthscale_{\pm} = \frac{\lamfac \lamupdn}{2 \pi} \, , \qquad
\timescale_{\pm} = \frac{\lamfac \lamupdn}{2 \pi \freqp \dupdn}
\end{align}
where $\lamfac$ is an integer to be chosen later. 
The explanation for these choices is as follows. First, we have chosen the characteristic amplitude, $\ampscale$, above to normalize the energy of the state-variable $u$ to unity, as will be demonstrated in Section \ref{tKdVSec}. Second, regarding $\lengthscale$, recall that $\lam$ is the characteristic wavelength corresponding to the peak forcing frequency $\freqp$ in the experiments. If only integer multiples of $\freqp$ were imposed (e.g.~lower frequencies were not present), then the forcing would produce waves that are periodic over lengthscale $\lam$. Since lower frequencies do exist, strict periodicity is not satisfied, but rather waves may be nearly periodic over the physical domain $\xi \in [-\lam/2, \lam/2]$. The approximation of near-periodicity becomes more accurate if integer multiples are considered, i.e.~$\xi \in [-\lamfac \lam/2, \lamfac \lam/2]$. Thus, we have chosen $\lengthscale$ above so that, over the dimensionless domain $\tilde{x} \in [-\pi, \pi]$, periodic boundary conditions can be  imposed on $u$ with an accuracy that increases with $\lamfac$.

	Lastly, regarding the characteristic timescale $\timescale$, the most basic timescale in the experiments is simply $\freqp^{-1}$, i.e.~the period of waves passing a fixed reference point. Of course, the leading-order behavior in shallow water is simple translation of waves with uniform speed $c$. The KdV equation provides the first correction to this behavior and describes dynamics that evolve over longer timescales. Hence we have rescaled $\freqp^{-1}$ by the factor $N \lam/(2 \pi \depth) \gg 1$, which provides a suitable long timescale in line with previous normalizations \cite{johnson1997modern}. The scales $\lengthscale = \lengthscale_{\pm}$ and $\timescale = \timescale_{\pm}$ change value across the ADC, which is important to note when comparing the theory against experimental measurements.
